---
title: "Team 12 - Hot Cocoa Codesters Project 1"
author: "Grace Chang, Seungwan Kim, Riley C Maher, Sage O'Toole, Jenna Kay Probst"
date: "2/21/2021"
output:
  html_document:
    number_sections: True
    toc: True
    toc_float: True
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r}
hotels <- read.csv("hotel_bookings.csv", stringsAsFactors = TRUE) #Reads in Hotel Bookings Data
library(ggplot2) #Calls ggplot
library(dplyr)
```
# General Analysis

## Arrival Distribution
```{r}
ggplot(hotels, aes(x = arrival_date_month)) +
  geom_bar(width = 0.8, fill="blue") +
  facet_wrap(~hotel) +
  labs(title = "Monthly Arrival Distribution") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) #Produces barplot of arrivals by month for each hotel

```

When looking towards the monthly distribution for both hotels, the City Hotel peaks in the summer and has lulls in late Fall and Winter. The Resort Hotel follows a similar, yet less volatile pattern. 

```{r}
hotels$seasons <- ifelse(hotels$arrival_date_month == "December"|
                                hotels$arrival_date_month == "January"|
                                hotels$arrival_date_month == "February",
                                "Winter",
                          ifelse(hotels$arrival_date_month == "March"|
                                 hotels$arrival_date_month == "April"|
                                 hotels$arrival_date_month == "May",
                                "Spring",
                          ifelse(hotels$arrival_date_month == "June"|
                                 hotels$arrival_date_month == "July"|
                                 hotels$arrival_date_month == "August",
                                "Summer",
                                "Autumn")))
sample_n(hotels, 10)
seasons <- table(hotels$season)
seasons <- round(prop.table(seasons),digits=3)

ggplot(hotels, aes(seasons)) + geom_bar(width = 0.8, fill="blue") +facet_wrap(~hotel) + labs(title = "Seasonal Arrival Distribution") #Produces barplot of arrivals by season for each hotel
```

This bar chart depicts another visual of the seasonal changes in arrivals for both hotels. Again, both hotels see the most arrivals in summer and the least in winter. (Approximately `r seasons[3]` of bookings in the Summer, and `r seasons[4]` in the Winter).The city hotel sees more drastic changes than the resort hotel as seen on the previous monthly distribution bar plot. 

```{r}
# Finding the types of hotels
hotel_types <- data.frame(table(hotels$hotel))
hotel_types <- hotel_types %>%
  rename(hotel_type = Var1, number_of_hotels = Freq)

# Creating a pie chart of hotel types
ggplot(hotel_types, aes(x="", y=number_of_hotels, fill=hotel_type)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar('y', start = 0) +
  geom_text(aes(y = number_of_hotels, label=number_of_hotels), position=position_stack(vjust=0.5), color = "black") +
  theme_void() +
  labs(title = "Number of Hotels by Type")
```

From this pie chart, we can see that there are `r hotel_types[1, 2]` City Hotels and `r hotel_types[2, 2]` Resort Hotels.

# Exploring Cancellation 
```{r hotel_cancellation}
total_cancel_rate <- round(NROW(hotels$is_canceled[hotels$is_canceled == 1])/NROW(hotels$is_canceled), 3)*100 #Returns proportion of bookings that were canceled
total_cancel_rate

cancel_prop <- round(prop.table(table(hotels$hotel, hotels$is_canceled),1)*100, 3) #Calculates cancellation rates for each hotel
cancel_prop

city_canceled <- filter(hotels, hotel == "City Hotel")
city_hotel <- ggplot(city_canceled, aes(x=is_canceled)) +
  geom_bar(aes(fill=is_canceled)) +
  scale_x_discrete(labels=c("Not Canceled", "Canceled")) +
  scale_fill_discrete(labels=c("Not Canceled", "Canceled")) +
  labs(y = "Number of Customers", x = "Bookings" , title = "City Hotels Booking Status") #Graph to show the city hotel cancellation
city_hotel

resort_canceled <- filter(hotels, hotel == "Resort Hotel")
resort_hotel <- ggplot(resort_canceled, aes(x=is_canceled)) +
  geom_bar(aes(fill=is_canceled) ) +
  scale_x_discrete(labels=c("Not Canceled", "Canceled")) +
  scale_fill_discrete(labels=c("Not Canceled", "Canceled")) +
  labs(y = "Number of Customers", x = "Bookings" , title = "Resort Hotels Booking Status") #Graph to show the resort hotel cancellation
resort_hotel
```

While the overall rate of cancellation is `r total_cancel_rate`%, the cancellation rate for City Hotels is slightly higher at `r cancel_prop[1, 2]`% and the cancellation rate for Resort Hotels is slightly lower at `r cancel_prop[1, 2]`%.
Customers who were planning to stay in the city hotels tended to cancel bookings more than those planning to stay in the resort hotels.

## Lead Time
```{r}
ggplot(hotels, aes(x=lead_time, fill= is_canceled)) +
  scale_fill_discrete(labels=c("Not Canceled", "Canceled")) +
  geom_histogram(binwidth = 50) #Dist of lead time separated by cancellation

ggplot(hotels, aes(x = lead_time, y = days_in_waiting_list)) +
  geom_point() # point plot to show the relationship between days in waiting list and lead time

lead_by_hotel <- ggplot(hotels, aes(x=hotel, y=lead_time))
lead_by_hotel + geom_boxplot() # Box plot to show the lead time by the hotel types
```
There does not appear to be a significant difference in the distribution of lead time for canceled and non-canceled bookings, with the exception of the jump in the cancellation rate from a lead time 0-50 to 50-100.

## Market Segmentation & Type and Distribution Channel
```{r}
round(prop.table(table(hotels$market_segment, hotels$is_canceled),1), 3) #Calculates cancellation rates by market segment 

round(prop.table(table(hotels$customer_type, hotels$is_canceled),1), 3) #Calculates cancellation by customer type

round(prop.table(table(hotels$distribution_channel,
                       hotels$is_canceled),1), 3) #Calculates cancellation rates by Channel

```
The cancellation rates based on market segment of the customer and the distribution channel the booking was made are quite interesting. For the market segment, the cancellation rates seem fairly consistent with the exception of Group bookings which have a cancellation rate nearly double the nearest segment (Online TA).Interestingly, when broken down into customer type, Group bookings have a much lower cancellation rate than the group market segment, a possible error is apparent in how the markets are categorized between these two measures (market segment vs customer type descriptions do not provide explanation in Kaggle). 

We see a similar scenario for the distribution sample, where corporate, direct, and GDS are all similar, but TA/TO (travel agents/tour operators) channel is nearly double that of any other channel excluding the undefined customers. 

## Deposit Type
```{r}
deposit_cancel <- round(prop.table(table(hotels$deposit_type,
                                         hotels$is_canceled),1), 3) #Calculates cancellation rates by deposit type
deposit_cancel[2,2]

plot(round(prop.table(table(hotels$deposit_type, hotels$is_canceled),1), 3),
     main="Cancellation by Deposit Type",
     ylab = "Cancellation",
     col=c("sky blue", "pink")) 
#Graphs cancellation rates by deposit type

```
The cancellation rates for each deposit type is incredible interesting. Among those with non-refundable deposits almost all of those customers ended up canceling their room, about `r deposit_cancel[2, 2]`. This made us doubt that the data was labeled incorrectly, but found in a Kaggle discussion post that the author of the paper, which the dataset is from, addressed this issue: "...through analysis of the “Nonrefundable” (DepositType) canceled bookings in some Asiatic countries (Country) and from certain distribution channels (DistributionChannel and Agent), it is possible to understand why so many “Nonrefundable” bookings are canceled. These bookings are usually made through OTA using false or invalid credit card details. These bookings are issued as support
for requests for visas to enter the country (a hotel booking is mandatory for applying for a Portuguese entry visa)." This is important information to consider when forecasting demand to make demand-management decisions.

## Past Customer Behavior
```{r}
repeat_cancel <- round(prop.table(table(hotels$is_repeated_guest, hotels$is_canceled),1), 3) #Calculates cancellation rates based on if they are repeat guest

round(prop.table(table(hotels$previous_cancellations, hotels$is_canceled),1), 3) #Calculates cancellation rates based on number of past canceled stays

table(hotels$previous_cancellations) #Calculates number of entries for each number of previous cancellations (Note that this isn't perfect since some cancellations likely happened within and outside of the cancellation period resulting in double counting of some people)
``` 
As expected the cancellation rate of customers who have previously stayed in the hotels is lower than those who haven't; by about half (`r repeat_cancel[1,2]` vs `r repeat_cancel[2,2]`).
Surprisingly, those who have had 1 previous cancellations have a higher rate of cancellation, but for those with 2-11 cancellations the rate is actually lower than people with no cancellations. It is quite interesting that those with more than 14 cancellations have a cancellation rate of 100%, you would not expect people to make this many bookings (up to 26) to cancel every single time.

## Exploring Arrival Times & Monthly cancellations
```{r}
ggplot(hotels, aes(x=arrival_date_day_of_month, fill =is_canceled)) +
  geom_histogram() +
  facet_wrap(~arrival_date_month) +
  scale_fill_discrete(labels=c("Not Canceled", "Canceled")) +
  labs(title = "Day of Arrival Distribution") #produces histograms of day of arrival for each month

month_cancel <- round(prop.table(table(hotels$arrival_date_month, hotels$is_canceled),1), 3) #Calculates cancellation rates by month

month_cancel
#Choose not to do distribution by year due to incomplete data for 2015 & 2017
```

The day of month distribution for each month is interesting, showing that in a majority of the months, arrivals peak between the tenth and twentieth. It should be noted that the separation of the cancellation shows that the cancellations follow the pattern of arrivals, as in when more people are set to arrive, more cancellations occur. The proportion table affirms this finding and is provided to show the exact range of cancellation rates for each month; with a range of `r month_cancel[5,2]` ~ `r month_cancel[7,2]` of bookings are canceled each month, there appears to be little variation among the rates per month. 

## Booking Changes
```{r}
ggplot(hotels, aes(x=booking_changes, fill = is_canceled)) +
  geom_histogram() +
  facet_wrap(~hotel) +
  coord_cartesian(xlim = c(0,6)) +
  scale_fill_discrete(labels=c("Not Canceled", "Canceled"))
```

# Exploring Booking Details

## Adults, Children, & Babies
```{r}
hotels$guests <- hotels$adults + hotels$children + hotels$babies #creates parameter of all guests on booking

ggplot(data=hotels, aes(adults)) +geom_histogram(binwidth = 1, fill="blue") + coord_cartesian(xlim = c(0,5))+facet_wrap(~hotel)
#Histogram of number adults for each hotel

ggplot(data=hotels, aes(children)) +geom_histogram(binwidth = 1, fill="red") + coord_cartesian(xlim = c(1,5), ylim = c(0,6000))+facet_wrap(~hotel)
#Histogram of number Children for each hotel

ggplot(data=hotels, aes(babies)) +geom_histogram(binwidth = 1, fill="red") + coord_cartesian(xlim = c(1,5), ylim = c(0,750))+facet_wrap(~hotel)
#Histogram of number babies for each hotel

ggplot(data=hotels, aes(guests)) +geom_histogram(binwidth = 1, fill="dark green") + coord_cartesian(xlim = c(0,7))+facet_wrap(~hotel)
#Histogram of total number of guests for each hotel
```
Above are histograms for the number of adults, children, babies, and all guests total for each hotel. Overwhelmingly the most common amount of adults is two, likely indicating that people typically travel as a couple or partnership. Since such a large portion of people travel without children or babies, the children and babies histogram start at 1. Most bookings with children only have 1, while the proportion of people have two children is much closer to one child for the resort hotel, possibly due to more people traveling to the resort for non-business purposes. For babies, it is overwhelmingly dominated by having only one baby for a booking. Besides the number of children distinction already mentioned, the distributions for each hotel are quite similar.

### Relationships between Adults & Children and Parking Spots Needed & Guests
```{r}
ggplot(data = hotels, aes(x=adults, y =children, color = "blue")) + geom_point(alpha=1/5) + coord_cartesian(xlim = c(0,10)) #creates scatter plot of number of adults and children on bookings

ggplot(data =hotels, aes(x=guests, y=required_car_parking_spaces, color = "blue")) + geom_point(alpha=1/5) #creates scatter plot of guests & number of parking spots required
```
Above the scatter plots are provided comparing the relationship between adults and children as well as the amount of guests on a booking and the amount of parking spots requested. As you can see, neither relationship appears to have significant patterns. It is interesting to note that for large amounts of guests, parking spots are not requested past ~8 guests. 

## Meals 
```{r}
# Seeing the type of meals families with kids buy
kids <- hotels %>%
  filter(children > 1)

meals_with_kids <- data.frame(table(kids$meal))

meals_with_kids <- meals_with_kids %>%
  mutate(meal_percent = round(Freq / nrow(kids), 3)*100) %>%
  rename(meal_type = Var1, number_of_meals = Freq)
meals_with_kids

ggplot(meals_with_kids, aes(x="", y=meal_percent, fill=meal_type)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar('y', start = 0) +
  geom_text(aes(y = meal_percent, label=meal_percent), position=position_stack(vjust=0.5), color = "black") +
  theme_void() +
  labs(title = "Perentage of Meals Bought by Adults with Kids")

# Seeing the type of meals families with no kids buy
no_kids <- hotels %>%
  filter(children == 0)

meals_with_no_kids <- data.frame(table(no_kids$meal))

meals_with_no_kids <- meals_with_no_kids %>%
  mutate(meal_percent = round(Freq / nrow(no_kids), 3)*100) %>%
  rename(meal_type = Var1, number_of_meals = Freq)
meals_with_no_kids

ggplot(meals_with_no_kids, aes(x="", y=meal_percent, fill=meal_type)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar('y', start = 0) +
  geom_text(aes(y = meal_percent, label=meal_percent), position=position_stack(vjust=0.5), color = "black") +
  theme_void() +
  labs(title = "Perentage of Meals Bought by Adults with No Kids")
```

From these pie charts we can see a notable difference between families who had kids and families who did not have kids when buying meals. The top meal choice of Bed & Breakfast did not change between groups, but the percentage bought did have a marked change. For the adults with kids, `r meals_with_kids[1, 3]`% bought the BB choice, while only `r meals_with_no_kids[1, 3]`% bought the BB choice. It seems that the adults with no kids group had a larger portion opt out of a meal choice (SC with `r meals_with_no_kids[4, 3]`% and Undefined with `r meals_with_no_kids[5, 3]`%) than the group of adults that had kids. This could show that the adults without kids are able to get meals in any means they see necessary, while the adults with kids would probably want convenience over everything while having more mouth to feed.

## Agent Bookings
```{r}
# Calculating the average daily rate of bookings that did/didn't use an agent
mean_adr_agent <- hotels %>%
  filter(agent != 'NULL') %>%
  summarize(mean_adr_agent = round(mean(adr, na.rm = TRUE), 2))

mean_adr_no_agent <- hotels %>%
  filter(agent == 'NULL') %>%
  summarize(mean_adr_no_agent = round(mean(adr, na.rm = TRUE), 2))

mean_adr_agent
mean_adr_no_agent
```

These numbers show us that, people booked with an agent actually paid more per night on average than those that did not book with an agent. If you used and agent, you would have paid \$`r mean_adr_agent[1]` while if you did not use an agent, you would have paid \$`r mean_adr_no_agent[1]`.

## Room Types
```{r}
table(hotels$reserved_room_type[hotels$hotel== "City Hotel"])
table(hotels$reserved_room_type[hotels$hotel== "Resort Hotel"])

hotels$reserved_room_type <- as.numeric(hotels$reserved_room_type)
hotels$assigned_room_type <- as.numeric(hotels$assigned_room_type)
hotels$RoomAquired <- ifelse(hotels$reserved_room_type == hotels$assigned_room_type,1,0)

acquired <- round(prop.table(table(hotels$hotel, hotels$RoomAquired), 1), 3)
acquired

```
As seen above, neither hotel has a 100% record for assigning guests to the rooms they reserved, with Resort Hotel having a slightly worse success rate of `r acquired[2,2]` vs `r acquired[1,2]`.

## Special Requests
```{r}
ggplot(data = hotels, aes(total_of_special_requests)) + geom_histogram(fill="blue") + facet_wrap(~hotel)
plot(hotels$total_of_special_requests,hotels$children)
plot(hotels$total_of_special_requests,hotels$babies)


```
Both hotels have similar distributions of special requests made, with most guests making 0 and a very small amount making more than 4. Neither amount of children or babies have a strong relationship with the amount of requests made. 


